---
title: "APD Crime Data"
author: "Rob Gordon"
date: '`r format(Sys.time(), "%b-%d-%Y")`'
output: 
  html_document: 
    toc: yes
    toc_depth: 4
---

Last Updated: `r format(Sys.time(), "%b-%d-%Y")`

## Introduction

Let's do stuff with the Atlanta Crime Data. We'll follow recommendations from [codeforatlanta/apd-crime-data github](https://github.com/codeforatlanta/apd-crime-data).

Namley:

- Get data
    - Download raw data file & unzip the csv contents.
    - Follow some of the logic from `getdata_APD_crime.R`
- This will output 2 data frames in the working directory:
    - `apdCrimeDataClean`: non-error rows, with extra columns X, X.1, X.2 removed
    - `apdCrimeDataErrors`: error rows only
    
The following packages must be loaded:

```{r, message=FALSE}
library(readxl)
library(tidyr)
library(dplyr)
library(leaflet)
library(Hmisc)
library(lubridate)
library(maptools)
library(foreign)
```
    
## Get Data

First we'll download the zip file. Commented lines were run once.

```{r}
# temp <- tempfile(tmpdir = "/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/")
# download.file(url = "http://www.atlantapd.org/files/crimedata/COBRA110416.zip"
              # ,destfile =  "/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/COBRA110416.zip")
apdCrimeData <- read_excel("/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/COBRA110416.xlsx"
  , sheet="Query")
# unlink(temp)
```

Let's check and make sure each column is what we expect it to be:

```{r}
apdCrimeData %>% str()
names(apdCrimeData) <- gsub(" ", "_", names(apdCrimeData))
apdCrimeDataTidy <- apdCrimeData %>% 
  mutate(MI_PRINX = as.numeric(MI_PRINX),
         offense_id = as.numeric(offense_id),
         rpt_date = lubridate::as_date(rpt_date, format = "%m/%d/%Y"),
         occur_date = lubridate::as_date(occur_date, format = "%m/%d/%Y"),
         poss_date = lubridate::as_date(poss_date, format = "%m/%d/%Y"),
         x = as.numeric(x),
         y = as.numeric(y))
```

Let's use pieces from `codeforatlanta`'s function to tidy this data:

```{r}
errors_horiz_offset <- c(91350923, 91420511, 91471067, 91521689, 101540909, 
                           101701138, 111971638, 112090917, 112411694, 113130827, 
                           113221244, 113270554, 113531411, 113590628, 120230979, 
                           122561142, 130101490, 141621526, 142570818, 151362710)
errors_strange_date <- c(141260924)
errors_all <- c(errors_horiz_offset, errors_strange_date)
apdCrimeDataClean <- apdCrimeDataTidy %>% 
  filter(!(offense_id %in% errors_all)) 
apdCrimeDataErrors <- apdCrimeDataTidy %>% 
  filter(offense_id %in% errors_all) 
```

Let's download the shapefiles directly from the atlanta crime data site:

```{r}
# temp <- tempfile(tmpdir = "/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/")
# download.file(url = "http://www.atlantapd.org/pdf/crime-data-downloads/1909FAB1-9E7F-4A34-8CDD-142D9DC83E7C.zip"
#               ,destfile =  "/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/1909FAB1-9E7F-4A34-8CDD-142D9DC83E7C.zip")
# unlink(temp)
beats <- readShapeSpatial("/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/1909FAB1-9E7F-4A34-8CDD-142D9DC83E7C/APD-Beats-070116_region.shp")
zones <- readShapeSpatial("/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/1909FAB1-9E7F-4A34-8CDD-142D9DC83E7C/APD-Zones-070116_region.shp")
```

Now let's add info from the .dbf files that came with the shapefiles.

```{r}
beats_dbf <- read.dbf("/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/1909FAB1-9E7F-4A34-8CDD-142D9DC83E7C/APD-Beats-070116_region.dbf")
zones_dbf <- read.dbf("/Users/robertgordon/Documents/apdCrimeData/apdFirstProj/apdCrime1/1909FAB1-9E7F-4A34-8CDD-142D9DC83E7C/APD-Zones-070116_region.dbf")
```

## Maps

Let's filter the data in different ways. `leaflet` seems unable to handle too many points on a map.

### First Map

Now let's plot 100 crimes on a map in leaflet.

```{r}
# describe(apdCrimeDataClean %>% select(x, y))
apdCrimeDataClean %>% 
  slice(1:100) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = ~x, lat = ~y, popup = ~UC2_Literal)
```

### Homicide map

How many crimes happened in 2016 alone?

```{r}
apdCrimeDataClean %>% 
  filter(year(occur_date) == 2016) %>% 
  nrow()
```

That might be too much for a single leaflet plot. What are the frequencies of all the different types of crime?

```{r}
apdCrimeDataClean %>% 
  group_by(UC2_Literal) %>% 
  summarise(freq = n()) %>% 
  ungroup()
```

Let's make a map of only homicides.

```{r}
apdCrimeDataClean %>% 
  filter(UC2_Literal=="HOMICIDE") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addCircles(lng = ~x, lat = ~y, popup = ~location, col="red")
```

### Add Shapefiles

Now we'll try to add shapefiles to our map. Let's first try adding the beats.

```{r}
apdCrimeDataClean %>% 
  filter(UC2_Literal=="HOMICIDE") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=beats, stroke=FALSE, fillOpacity = 0.5, smoothFactor=0.5,
              color="gray") %>% 
  addPolylines(data=beats, stroke=TRUE, fillOpacity = 0.5, smoothFactor=0.5,
              color="black") %>%
  addCircles(lng = ~x, lat = ~y, popup = ~location, col="red")
```

Now add zones.

```{r}
apdCrimeDataClean %>% 
  filter(UC2_Literal=="HOMICIDE") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=zones, stroke=FALSE, fillOpacity = 0.5, smoothFactor=0.5,
              color="gray") %>% 
  addPolylines(data=zones, stroke=TRUE, fillOpacity = 0.5, smoothFactor=0.5,
              color="black") %>%
  addCircles(lng = ~x, lat = ~y, popup = ~location, col="red")
```

What if I add both zones and beats?

```{r}
apdCrimeDataClean %>% 
  filter(UC2_Literal=="HOMICIDE") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=zones, stroke=FALSE, fillOpacity = 0.5, smoothFactor=0.5,
              color="gray") %>% 
  addPolylines(data=zones, stroke=TRUE, fillOpacity = 0.5, smoothFactor=0.5,
              color="blue") %>%
  addPolylines(data=beats, stroke=TRUE, fillOpacity = 0.5, smoothFactor=0.5,
              color="red") %>%
  addCircles(lng = ~x, lat = ~y, popup = ~location, col="red")
```

## Color by Beat

Can we color the beats or zones by total number of crime type?

### Homicide

```{r}
homicidesByBeat <- apdCrimeDataClean %>% 
  filter(UC2_Literal=="HOMICIDE") %>% 
  group_by(beat) %>% 
  summarise(freq = n()) %>% 
  ungroup()
# heatCols <- heat.colors(nrow(homicidesByBeat))[cut(sort(homicidesByBeat$freq),nrow(homicidesByBeat))]
pal <- colorNumeric(
  # palette = "YlGnBu",
  # palette = "RdYlBu",
  palette = "YlOrRd",
  domain = homicidesByBeat$freq
)
apdCrimeDataClean %>% 
  # filter(UC2_Literal=="HOMICIDE") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolylines(data=beats, stroke=TRUE, fillOpacity = 0.5, smoothFactor=0.5,
              color="black") %>%
  # addPolygons(data=beats, stroke=FALSE, fillOpacity = 0.5, smoothFactor=0.5,
              # fillColor=heatCols) %>% 
  addPolygons(data=beats, stroke=FALSE, fillOpacity = 0.5, smoothFactor = 0.5,
              color = ~pal(homicidesByBeat$freq)) %>% 
  addLegend("bottomright", pal = pal, values = homicidesByBeat$freq,
            title = "Legend",
            opacity = 1)
```

### Rape

```{r}
rapeByBeat <- apdCrimeDataClean %>% 
  filter(UC2_Literal=="RAPE") %>% 
  group_by(beat) %>% 
  summarise(freq = n()) %>% 
  ungroup()
# heatCols <- heat.colors(nrow(rapeByBeat))[cut(sort(rapeByBeat$freq),nrow(rapeByBeat))]
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = rapeByBeat$freq
)
apdCrimeDataClean %>% 
  # filter(UC2_Literal=="RAPE") %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolylines(data=beats, stroke=TRUE, fillOpacity = 0.5, smoothFactor=0.5,
              color="black") %>%
  # addPolygons(data=beats, stroke=FALSE, fillOpacity = 0.5, smoothFactor=0.5,
              # fillColor=heatCols) %>% 
  addPolygons(data=beats, stroke=FALSE, fillOpacity = 0.5, smoothFactor = 0.5,
              color = ~pal(rapeByBeat$freq)) %>% 
  addLegend("bottomright", pal = pal, values = rapeByBeat$freq,
            title = "Legend",
            opacity = 1)
```

## Session Info

```{r}
sessionInfo()
```
